Multi-Core Model Checking and Maximum Satisfiability Applied to Hardware-Software Partitioning

### 2016-08-09 Docs/IJES/revised/journal.tex
revised paper at 8th of August by ABT

Optimization is the act of obtaining the best result ({\it i.e.}, the optimal solution) under given circumstances as defined in~\cite{Rao2009}. There is no single method available for efficiently solving all optimization problems, according to~\cite{Rao2009}. The most well-known technique is linear programming, which is an method applicable for the solution of problems in which the objective function and the constraints appear as linear functions of the decision variables. A particular case of linear programming is ILP, in which the variables can assume just integer values. Eq.~\ref{linear-programming-problem} shows a typical linear programming problem, where $A$ and $b$ are vectors or matrixes that describe the constraints

@@

\begin{itemize}
\item{\textbf{maximize(T)}
this function informs to the solver that a given variable $T$ should be maximized, which includes real, integer, or bit-vector variables.}
\item{\textbf{minimize(T)}
this function informs the solver that a given variable $T$ should be minimized, the accepted types are the same as maximize function.}
\item{\textbf{Assert-Soft F : weight n}
the function \textit{assert-soft} adds a restriction to $F$, which can also add a weight $n$; the default value is $1$.}
\end{itemize}

@@

The inputs of the problem are: a directed simple graph $ G = (V,E) $, called the task graph of the system, is necessary. The vertices $V = \{V_1,V_2,\dotso,V_n\}$ represent the nodes that are the components of the system that will be partitioned. The edges $E$ represent communication between components. Additionally, each node  $V_i$ has a cost $h(V_i)$ (or $h_i$) of hardware if implemented in hardware and a cost $s(V_i)$ (or $ s_i $) of software if implemented in software. Finally, $c(V_i,V_j)$ represents the communication cost between $V_i$ and $V_j$ if they are implemented in different contexts (hardware or software).

Based on~\cite{Arato2003}, is called a hardware-software partition if it is a bipartition of $V:P = (V_h, V_s)$, where $V_h \cup V_s = V$  and $V_h \cap V_s = 0$. The crossing edges are $E_p = \{(V_i,V_j):V_i \in V_s, V_j \in V_h$ or $V_i \in V_h, V_j \in V_s $. The hardware cost of $P$ is given by Eq.~\ref{hardware-costs}

@@

\vspace{2 mm}
Three different optimization and decision problems can be defined. In this paper, the focus is on the case that $ S_0 $ is given, {\it i.e.}, to find a $P$ HW-SW partitioning so that $ S_p \leq S_0 $ and $ H_p $ is minimal, which is thus related to system with hard real-time constraints. Based on Eq.~\ref{linear-programming-problem} and~\ref{software-communication-costs}, the constraints can be reformulated as

@@

\vspace{2 mm}
\noindent where $x$ represents the decision variable. Concerning the complexity of this problem, \cite{Arato2003} demonstrate that it is NP-Hard, as defined in~\cite{Cormem}.

@@

The ILP and GA were taken from our previous studies, from~\cite{Trindade2015} and~\cite{Trindade2016}. Both use slack variables in order to be possible to represent constraints and to use commercial tools. However, GA had improvements from the parameters of related studies in order to increase the solution accuracy without producing timeout. The tuning was performed by empirical tests and resulted in changing of three parameters, which are passed to the function \textit{ga} of MATLAB, as shown in~\cite{TheMathWorks2013}: the population size was set from $300$ to $500$, the Elite count changed from $2$ (default value) to $50$, and the number of generations changed from $100*NumberOfVariables$ (default) to $75$.

@@

The ESBMC algorithm starts with the declarations of hardware, software, and communication costs. $S_0$ must also be defined, as the transposed incidence matrix and the identity matrix, as typically done in MATLAB. Here, matrices $A$ and $b$ are generated. At that point, the ESBMC algorithm starts to differ from the ILP and GA presented in~\cite{Trindade2015}.

\begin{lstlisting}[basicstyle=\footnotesize,caption={Pseudocode describing sequential ESBMC.},label={Pseudocode-describing-ESBMC},numbersep=7pt,frame=tb,captionpos=t,numberstyle=\tiny,mathescape=true]
Initialize variables 
Declare number of nodes and edges
Declare hardware cost of each node as array (h)
Declare software cost of each node as array (s)
Declare communication cost of each edge (c)
Declare the initial sw cost of ($S_0$)
Declare transposed incidence matrix graph G(E)
Define the solutions ($X_i$) as Boolean
main {
 For TipH = 0 to Hmax Do {
  Populate $X_i$ with nondet / test values
  Calculate s(1-x)+c|Ex| and store at variable
  Requirement isued by Assume (variable <= $S_0$)
  Calculate Hp cost Based on value of $X_i$
  Violation check with Assert(Hp > TipH)
 }
}
\end{lstlisting}

\vspace{2 mm}
It is possible to inform to ESBMC with which type of values the variables must be tested. Therefore, there is a declaration to populate all decision variables $x$ with non-deterministic Boolean values. Those values that change for each test will generate a possible solution and obey the constraints. If this is achieved, then a feasible solution is found and the ASSUME directive is responsible for ensuring the compliance of those constrains (\textit{i.e.}, $A.x \leq b$).

A loop controls the cost of hardware hint, starting with zero and reaching the maximum value considering the case, where all nodes are partitioned to hardware. To every test performed, the hardware hint is compared to the feasible solution. This is accomplished by an \textit{ASSERT} statement at the end of the algorithm, a predicate that controls the halt condition (a \textit{true-false} statement). If the predicate is \textit{FALSE}, then the optimization is finished, {\it i.e.}, the solution is found.

@@

In the ESBMC algorithm is not necessary to add slack variables, because the modulus operation is kept, which reduces the number of variables to be solved.

@@

The arithmetic expressions from lines $10$ to $12$ represent the constraints described in Eq.~\ref{hw-sw-partitioning}. Here, variable \textit{SC} refers to the software cost, while \textit{CC} denotes the communication cost. In line $12$, the \textit{Fobj} (objective function) is declared, which denotes the product between the hardware cost and the decision variables vector, which contains only Boolean values. \textit{Fobj} should be minimized to obtain the optimal hardware solution. To achieve this, two constraints are imposed to ESBMC-$\nu$Z: the first one refers to the sum of the software and communication costs, where the result should be less than $S_0$; and the second one informs to ESBMC-$\nu$Z that \textit{Fobj} should be minimized. Finally, the model is checked by ESBMC-$\nu$Z and if there is a solution that 
meets the constraints, then the \textit{Fobj} value is provided.

@@

\bibitem[\protect\citeauthoryear{Paci et al}{2008}]{Federica2008}
Paci, F., Ferrini, R., Sun, Y. and Bertino, E. (2008) 'Authorization and User Failure Resiliency for WS-BPEL business processes', {\it Proceedings of the International Conference on Service-Oriented Computing}, pp.116--131

>>>>

Optimization is the act of obtaining the best result ({\it i.e.}, the optimal solution) under given circumstances as defined in~\cite{Rao2009}. There is no single method available for efficiently solving all optimization problems, according to~\cite{Rao2009}. The most well-known technique is linear programming, which is an method applicable for the solution of problems in which the objective function and the constraints appear as linear functions of the decision variables. A particular case of linear programming is ILP, in which the variables can assume just integer values. Eq.~\ref{linear-programming-problem} shows a typical linear programming problem, where $A$ and $b$ are vectors from the objective function, and $Aeq$ and $beq$ are matrixes that describe the linear equality constraints

@@

\begin{itemize}
\item{\textbf{maximize(T)}
this function instructs the solver that a given variable $T$ should be maximized, which includes real, integer, or bit-vector variables.}
\item{\textbf{minimize(T)}
this function instructs the solver that a given variable $T$ should be minimized, the accepted types are the same as maximize function.}
\item{\textbf{Assert-Soft F : weight n}
the function \textit{assert-soft} adds a restriction to $F$, which can also add a weight $n$; the default value is $1$.}
\end{itemize}

@@

A directed simple graph $ G = (V,E) $, called the task graph of the system, is given. Where the vertices $V = \{V_1,V_2,\dotso,V_n\}$ represent the nodes that are the components of the system that will be partitioned. The edges $E$ represent communication between components. Additionally, each node  $V_i$ has a cost $h(V_i)$ (or $h_i$) of hardware if implemented in hardware and a cost $s(V_i)$ (or $ s_i $) of software if implemented in software. Finally, $c(V_i,V_j)$ represents the communication cost between $V_i$ and $V_j$ if they are implemented in different contexts (hardware or software).

$P$ is called a hardware-software partition if it is a bipartition of $V:P = (V_h, V_s)$, where $V_h \cup V_s = V$  and $V_h \cap V_s = 0$. The crossing edges are $E_p = \{(V_i,V_j):V_i \in V_s, V_j \in V_h$ or $V_i \in V_h, V_j \in V_s $, \cite{Arato2003}. The hardware cost of $P$ is given by Eq.~\ref{hardware-costs}

@@

\vspace{2 mm}
Different optimization (and decision) problems can be defined, \cite{Arato2003}. In this paper, the focus is on systems with hard real-time constraints: $ S_0 $ is given (initial cost of software), {\it i.e.}, the goal is to find a $P$ HW-SW partitioning so that $ S_p \leq S_0 $ and $ H_p $ is minimal, which is thus related . Based on Eq.~\ref{linear-programming-problem} and~\ref{software-communication-costs}, the constraints can be reformulated as

@@

\vspace{2 mm}
\noindent where $s$ and $c$ are the vectors representing the cost functions, $E$ is the transposed incidence matrix of $G$ (indicating which edges cross the boundary between the contexts of hardware and software), and $x$ represents the decision variable (a binary vector indicating the partition: 1 if the node is realized in hardware and 0 if node is realized in software). Concerning the complexity of this problem, \cite{Arato2003} demonstrate that it is NP-Hard, as defined in~\cite{Cormem}.

@@

The ILP and GA were taken from our previous studies, from~\cite{Trindade2015} and~\cite{Trindade2016}. Both use slack variables in order to eliminate the modulus operator of Eq.~\ref{hw-sw-partitioning} and to use commercial tools. However, GA had improvements from the parameters of related studies in order to increase the solution accuracy without producing timeout. The tuning was performed by empirical tests and resulted in changing of three parameters, which are passed to the function \textit{ga} of MATLAB, as shown in~\cite{TheMathWorks2013}: the population size was set from $300$ to $500$, the Elite count changed from $2$ (default value) to $50$, and the number of generations changed from $100*NumberOfVariables$ (default) to $75$.

@@

The ESBMC algorithm starts with the declarations of hardware, software, and communication costs. $S_0$ must also be defined, as the transposed incidence matrix (used in Eq.~\ref{hw-sw-partitioning}) and the identity matrix (necessary to work with the matrices), as typically done in MATLAB. Here, matrices $A$ and $b$ are generated. At that point, the ESBMC algorithm starts to differ from the ILP and GA presented in~\cite{Trindade2015}.

\begin{lstlisting}[basicstyle=\footnotesize,caption={Pseudocode describing sequential ESBMC.},label={Pseudocode-describing-ESBMC},numbersep=7pt,frame=tb,captionpos=t,numberstyle=\tiny,mathescape=true]
Initialize variables 
Declare number of nodes and edges
Declare hardware cost of each node as array (h)
Declare software cost of each node as array (s)
Declare communication cost of each edge (c)
Declare the initial sw cost of ($S_0$)
Declare transposed incidence matrix graph G(E)
Define the solutions ($x_i$) as Boolean
main {
 For TipH = 0 to Hmax Do {
  Populate $x_i$ with nondet (the values of test)
  Calculate s_i(1-x_i)+c|Ex_i| and store at variable
  Requirement enforced by Assume (variable <= $S_0$)
  Calculate Hp cost Based on value of $x_i$
  Violation check with Assert(Hp > TipH)
 }
}
\end{lstlisting}

\vspace{2 mm}
It is possible to instruct ESBMC with which type of values the variables must be tested. Therefore, there is a declaration to populate all decision variables $x$ with non-deterministic Boolean values. Those values that change for each test will generate a possible solution and obey the constraints. If this is achieved, then a feasible solution is found and the ASSUME directive is responsible for ensuring the compliance of those constrains (\textit{i.e.}, $A.x \leq b$).

A loop controls the cost of hardware hint, starting with zero and reaching the maximum value considering the case, where all nodes are partitioned to hardware (Hmax). To every test performed, the hardware hint is compared to the feasible solution. This is accomplished by an \textit{ASSERT} statement at the end of the algorithm, a predicate that controls the halt condition (a \textit{true-false} statement). If the predicate is \textit{FALSE}, then the optimization is finished, {\it i.e.}, the solution is found.

@@

In the ESBMC algorithm is not necessary to add slack variables at Eq.~\ref{hw-sw-partitioning}, which reduces the number of variables to be solved when compared to ILP and GA.

@@

The arithmetic expressions from lines $10$ to $12$ represent the constraints described in Eq.~\ref{hw-sw-partitioning}. Here, variable \textit{SC} refers to the software cost, while \textit{CC} denotes the communication cost. In line $12$, the \textit{Fobj} (objective function) is declared, which denotes the product between the hardware cost and the decision variables vector, which contains only Boolean values. \textit{Fobj} should be minimized to obtain the optimal hardware solution. To achieve this, two constraints are imposed to ESBMC-$\nu$Z: the first one refers to the sum of the software and communication costs, where the result should be less than $S_0$; and the second one instructs to ESBMC-$\nu$Z that \textit{Fobj} should be minimized. Finally, the model is checked by ESBMC-$\nu$Z and if there is a solution that 
meets the constraints, then the \textit{Fobj} value is provided.

@@

\bibitem[\protect\citeauthoryear{Paci \textit{et al.}}{2008}]{Federica2008}
Paci, F., Ferrini, R., Sun, Y. and Bertino, E. (2008) 'Authorization and User Failure Resiliency for WS-BPEL business processes', {\it Proceedings of the International Conference on Service-Oriented Computing}, pp.116--131
